Probability

### 2016-04-22 prob.tex
plan

In the remainder of the paper, we consider variations of quantum
probability spaces motivated by computation of numerical quantities in
a world with limited resources:
\begin{itemize}
\item Instead of the Hilbert space $\Hilb$ (constructed over the
  uncountable and uncomputable complex numbers $\mathbb{C}$), we will
  consider variants constructed over finite
  fields~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.
\item Instead of real-valued probability measures producing results in
  the uncountable and uncomputable interval $[0,1]$, we will consider
  set-valued probablitity measures~\cite{Artstein1972,PuriRalescu1983}.
\end{itemize}
We will then ask if it is possible to construct variants of quantum
probability spaces under these conditions. The main question is
related to the definition of probability measures: is it possible to
still define a probability measure as a function that depends on a
single state? Specifically,
\begin{itemize}
\item given a state $\ket{\psi}$, is there a probability measure
  mapping events to probabilities that only depends on $\ket{\psi}$?
  In the conventional quantum probability space, the answer is yes by
  the Born rule~\cite{Born1984,Mermin2007} and the map is given by:
  $E \mapsto \ip{\psi}{E\psi}$.
\item given a probability measure $\mu$
  mapping each event $E$
  to a probability, is there is a \emph{unique} state $\psi$
  such that $\mu(E)
  =
  \ip{\psi}{E\psi}$. In the conventional case, the answer is yes by
  Gleason's
  theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.
\end{itemize}

\amr{The rest is NO NO NO. We are inventing some other conditions on
  probability measures that are not part of the framework !!! }

If there may be more than one probability measure, we will discuss
whether we will keep using the Born rule (\ref{eq:Born}) or there
is another formula $\tilde{\mu}$ such that $\tilde{\mu}_{\psi}$
is a probability measure for all $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $.

Then, Gleason's theorem states given a probability measure $\mu$
there is a mixed state $\ket{\psi}$ such that $\mu=\mu_{\psi}$,
i.e., the Born rule is surjective\footnote{If we extend the domain of $\mu_{\psi}$ including the mixed states.}.
For any other formula~$\tilde{\mu}$, we can ask whether $\tilde{\mu}$
is surjective as well.

Obveriously, we don't want arbirarily assign a state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
with a proabability measure $\tilde{\mu}_{\psi}$, for example, assigning
every state to the same probability measure. We want $\tilde{\mu}_{\psi}$
satisfying the following properties with some physical meaning:
\begin{itemize}
\item $\ip{\psi}{\phi}=0\Leftrightarrow\tilde{\mu}_{\psi}\left(\proj{\phi}\right)=\tilde{0}$,
where $\tilde{0}$ is $0$ for $\left[0,1\right]$ and $\tilde{0}$
is impossible for $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $. 
\item $\tilde{\mu}_{\psi}\left(\proj{\phi}\right)=\tilde{\mu}_{\mathbf{U}\ket{\psi}}\left(\mathbf{U}\proj{\phi}\mathbf{U}^{\dagger}\right)\textrm{ ,}$where
$\ket{\psi},\ket{\phi}$ are states and $\mathbf{U}$ is any unitary
map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbb{1}$. 
\end{itemize}
And the results can be summarized in the following table:\\
\center{ %
\begin{tabular}{>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}}
\hline 
State space $\Hilb$  & Probability values  & Is there a $\tilde{\mu}$ satisfying the given conditions?  & Is the $\tilde{\mu}$ surjective?\tabularnewline
\hline 
\hline 
$\C^{d}$ for $d\ge3$  & $\left[0,1\right]$  & Yes  & Yes\tabularnewline
\hline 
$\C^{d}$  & $\mathscr{L}_{2}$  & Yes  & No\tabularnewline
\hline 
$\ffzd{p^{2}}$for $d\ge3$ except $d=p=3$  & $\left[0,1\right]$  & No  & \tabularnewline
\hline 
$\ffzd{p^{2}}$  & $\mathscr{L}_{2}$  & Yes & No\tabularnewline
\hline 
\end{tabular}}

>>>>

In the remainder of the paper, we consider variations of quantum
probability spaces motivated by computation of numerical quantities in
a world with limited resources:
\begin{itemize}
\item Instead of the Hilbert space $\Hilb$ (constructed over the
  uncountable and uncomputable complex numbers $\mathbb{C}$), we will
  consider variants constructed over finite
  fields~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.
\item Instead of real-valued probability measures producing results in
  the uncountable and uncomputable interval $[0,1]$, we will consider
  finite set-valued probablitity measures~\cite{Artstein1972,PuriRalescu1983}.
\end{itemize}
We will then ask if it is possible to construct variants of quantum
probability spaces under these conditions. The main question is
related to the definition of probability measures: is it possible to
still define a probability measure as a function that depends on a
single state? Specifically,
\begin{itemize}
\item given a state $\ket{\psi}$, is there a probability measure
  mapping events to probabilities that only depends on $\ket{\psi}$?
  In the conventional quantum probability space, the answer is yes by
  the Born rule~\cite{Born1984,Mermin2007} and the map is given by:
  $E \mapsto \ip{\psi}{E\psi}$.
\item given a probability measure $\mu$
  mapping each event $E$
  to a probability, is there is a \emph{unique} state $\psi$
  such that $\mu(E)
  =
  \ip{\psi}{E\psi}$. In the conventional case, the answer is yes by
  Gleason's
  theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.
\end{itemize}

% If there may be more than one probability measure, we will discuss
% whether we will keep using the Born rule (\ref{eq:Born}) or there
% is another formula $\tilde{\mu}$ such that $\tilde{\mu}_{\psi}$
% is a probability measure for all $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $.

% Then, Gleason's theorem states given a probability measure $\mu$
% there is a mixed state $\ket{\psi}$ such that $\mu=\mu_{\psi}$,
% i.e., the Born rule is surjective\footnote{If we extend the domain of $\mu_{\psi}$ including the mixed states.}.
% For any other formula~$\tilde{\mu}$, we can ask whether $\tilde{\mu}$
% is surjective as well.

% Obveriously, we don't want arbirarily assign a state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
% with a proabability measure $\tilde{\mu}_{\psi}$, for example, assigning
% every state to the same probability measure. We want $\tilde{\mu}_{\psi}$
% satisfying the following properties with some physical meaning:
% \begin{itemize}
% \item $\ip{\psi}{\phi}=0\Leftrightarrow\tilde{\mu}_{\psi}\left(\proj{\phi}\right)=\tilde{0}$,
% where $\tilde{0}$ is $0$ for $\left[0,1\right]$ and $\tilde{0}$
% is impossible for $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $. 
% \item $\tilde{\mu}_{\psi}\left(\proj{\phi}\right)=\tilde{\mu}_{\mathbf{U}\ket{\psi}}\left(\mathbf{U}\proj{\phi}\mathbf{U}^{\dagger}\right)\textrm{ ,}$where
% $\ket{\psi},\ket{\phi}$ are states and $\mathbf{U}$ is any unitary
% map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbb{1}$. 
% \end{itemize}
% And the results can be summarized in the following table:\\
% \center{ %
% \begin{tabular}{>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}}
% \hline 
% State space $\Hilb$  & Probability values  & Is there a $\tilde{\mu}$ satisfying the given conditions?  & Is the $\tilde{\mu}$ surjective?\tabularnewline
% \hline 
% \hline 
% $\C^{d}$ for $d\ge3$  & $\left[0,1\right]$  & Yes  & Yes\tabularnewline
% \hline 
% $\C^{d}$  & $\mathscr{L}_{2}$  & Yes  & No\tabularnewline
% \hline 
% $\ffzd{p^{2}}$for $d\ge3$ except $d=p=3$  & $\left[0,1\right]$  & No  & \tabularnewline
% \hline 
% $\ffzd{p^{2}}$  & $\mathscr{L}_{2}$  & Yes & No\tabularnewline
% \hline 
% \end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{All Continuous or All Discrete}

Before we turn to the main part of the paper, we quickly dismiss the
possibility of having one but not the other of the discrete
variations. Speficially, it is impossible to maintain the Hllbert
space and have a finite set-valued probability measure and it is also
impossible to have a vector space constructed over a finite field with
a real-valued probability measure.

%%%
\subsection{HIlbert Space with Finite Set-Valued Probability Measure}

%%%
\subsection{Discrete Vector Space with Real-Valued Probability Measure}
